---
title: "STA-380-Homework-1-Dalton-Haines-Jensen-Karnes"
output: html_document
---
Contributors: Reed Dalton, Corey Haines, Alex Jansen, McKenna Karnes

#Probability Practice

### Part A. 
Given a single survey question to visitors to a particular website, what fraction of people who truthfully answered the question answered yes?      
* There are two types of clickers: Random Clickers (RC), and Truthful Clickers (TC)     
* There are two answers to the survey: yes and no   
* The expected fraction of Random Clickers (RC) is 0.3     
* The expected fraction of Truthful Clickers (TC) is 0.7     
* The probability that the Random Clickers (RC) selects yes is .5   
* The probability that the Random Clickers (RC) selects no is .5      
The results of the survey were such that 65% said Yes and 35% said No.          
```{r}
prob.yes = 0.65
prob.no = 0.35
prob.yes.given.RC = 0.5
prob.no.given.RC = 0.5
prob.TC = 0.7
prob.RC = 0.3
prob.yes.given.TC = (prob.yes - (prob.RC * prob.yes.given.RC))/prob.TC

```
`prob.yes` = `prob.TC` \* `prob.yes.given.TC` + `prob.RC` \* `prob.yes.given.RC`

The percentage of Truthful Clickers who answered "Yes" to the the simple survey is `r prob.yes.given.TC`

### Part B.
Medical test for disease using Bayes Formula. 
* The sensitivity is about 0.993. That is, if someone has the disease, there is a probability of 0.993 that they will test positive.     
* The specificity is about 0.9999. This means that if someone doesn't have the disease, there is probability of 0.9999 that they will test negative.     
* About 0.0025% of all people in the general population have the disease
If someone tests positive for the disease, what is the probability they actually have the disease?

```{r}
pop.infected = 0.000025
pop.clean = 0.999975

infected.pos = 0.993
infected.neg = 0.007

clean.pos = 0.0001
clean.neg = 0.9999
infected.test.pos = (infected.pos*pop.infected)/(infected.pos*pop.infected + clean.pos*pop.clean)
infected.test.pos = infected.test.pos * 100
infected.test.pos
```
If someone tested positive for the disease, there is only a `r infected.test.pos`% chance that they actually have the disease. 
Given our calculations, implementing a universal testing policy for this disease would be problematic because over 80% of the time a test for the disease is positive, the person will not actually be infected. This could lead to mass hysteria and put unneeded strain on the population's health systems. 
=======
prob.yes == prob.TC * prob.yes.given.TC + prob.RC * prob.yes.given.RC
prob.yes.given.TC
```
>>>>>>> 7521a9699de034e26960f9f3df5c0979caeec7e1
